The Form of Uncertainty Affects Selection for Social Learning

### 2023-02-18 ms.tex
updates for revisions from local desktop

\usepackage{booktabs}
\usepackage{pslatex}
% \usepackage{apacite}
\usepackage{caption}
\usepackage{natbib}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{setspace}
\doublespace
% \usepackage{url}
\usepackage{bigfoot}
\usepackage[export]{adjustbox}
\setlength\intextsep{0pt}

@@

\paragraph{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviours
\citep{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between these payoffs is usually taken to influence the strength of selection on learning strategies.
However, the ability to discern payoff differences between behaviours is also a source of uncertainty. In reality, payoffs for particular behaviours are not always consistent. A behaviour may yield a large payoff sometimes and a small
payoff other times \citep{McElreath2005}. This means that signals about the
relationship between behaviour and payoff are often noisy, and differentiating
between behavioural options is in part a problem of signal detection. When the
difference in expected payoffs between optimal and non-optimal behaviours is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviours are similar relative to the size of
their variances, ambiguity arises about which behaviours really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\citep{perreault2012bayesian}
and empirical~\citep{McElreath2005,Toyokawa2019} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection-set size in that it not only 
affects both uncertainty \emph{and} the strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. 
\edit{It is a major problem to understand how payoff variance and payoff distributions
generally affect evolved behavioural strategies~\citep{Haaland2019}. For the sake
of theoretical clarity considering the \emph{interaction} of various sources of uncertainty
we only operationalize payoff ambiguity as the difference between expected 
payoffs, even though empirical studies often include both forms of payoff
uncertainty~\citep{McElreath2005,Toyokawa2019}, and a more complete
model would include differences in mean and variance of payoffs.}

>>>>

\usepackage{booktabs}
\usepackage{pslatex}
\usepackage{caption}
\usepackage{natbib}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{setspace}
\doublespace
% \usepackage{url}
\usepackage{bigfoot}
\usepackage[export]{adjustbox}
\setlength\intextsep{0pt}

@@

\paragraph{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviours
\citep{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between these payoffs is usually taken to influence the strength of selection on learning strategies.
However, the ability to discern payoff differences between behaviours is also a source of uncertainty. In reality, payoffs for particular behaviours are not always consistent. A behaviour may yield a large payoff sometimes and a small
payoff other times \citep{McElreath2005}. This means that signals about the
relationship between behaviour and payoff are often noisy, and differentiating
between behavioural options is in part a problem of signal detection. When the
difference in expected payoffs between optimal and non-optimal behaviours is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviours are similar relative to the size of
their variances, ambiguity arises about which behaviours really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\citep{perreault2012bayesian}
and empirical~\citep{McElreath2005,Toyokawa2019} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection-set size in that it not only 
affects both uncertainty \emph{and} the strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. 
\edit{It is an outstanding problem to understand how payoff variance and payoff distributions
generally affect evolved behavioural strategies~\citep{Haaland2019}. For the sake
of theoretical clarity for our goal of understanding the \emph{interaction} of various sources of uncertainty
we only operationalize payoff ambiguity as the difference between expected 
payoffs, even though empirical studies often include both forms of payoff
uncertainty~\citep{McElreath2005,Toyokawa2019}, and a more complete
model would include differences in mean and variance of payoffs.}
